---
title: "Factors and Happiness"
author: "Yunlin Zhou"
date: "12/9/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
```{r echo = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(jtools)
library(patchwork)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
knitr::opts_chunk$set(warning=FALSE, message = FALSE)


```



<br>

### First sight on factors & happiness score

To better explore the relationship between happiness score and every factor, we made the linear regression plots and calculated related estimates. Further, we analyzed the result and derived to a more direct and accurate conclusion. 

```{r}
# Data Preparation
round2 = function(x){
  if(is.numeric(x))
  x= round(x, digits = 2)
  return(x)}

happy_df_fac = readxl::read_xls('DataPanelWHR2021C2.xls') %>% 
  janitor::clean_names() %>% 
  filter(year < 2019) %>% 
  map_dfc(round2) %>%
  mutate(year = as.character(year)) %>% 
  rename("happiness_score"= life_ladder)



# Graphic function
gg_factor = function(A,B,name){
  name = name %>% 
    str_replace_all("_"," ") %>% 
    str_to_title()
  
  ggplot(happy_df_fac, aes(x = A, y = B))+
    geom_point(aes(color = year)) +
    geom_smooth(method = "lm") +
    labs(
    x = name,
    y = "Happiness Score",
  )
}
```

```{r fig.height = 15, fig.width = 14}
# Plot

gg_log = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$log_gdp_per_capita, name = "log_gdp_per_capita")
gg_social = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$social_support, name = "social_support")
gg_healthy = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$healthy_life_expectancy_at_birth, name = "healthy_life_expectancy_at_birth")
gg_freedom = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$freedom_to_make_life_choices, name = "freedom_to_make_life_choices")
gg_generosity = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$generosity, name = "generosity")
gg_perceptions = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$perceptions_of_corruption, name = "perceptions_of_corruption")
gg_positive = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$positive_affect, name = "positive_affect")
gg_negative = gg_factor(A=happy_df_fac$happiness_score,B =happy_df_fac$negative_affect, name = "negative_affect")

library(patchwork)
(gg_log + gg_social+ gg_healthy + gg_freedom)/(gg_generosity + gg_perceptions+ gg_positive + gg_negative)+ 
    plot_layout(guides = "collect") + 
    plot_annotation(title = "First sight on factors & happiness score") & 
    theme(
      legend.position = "bottom")

```

As we can see in the plots, all the factors are related to the happiness score. People would feel happy if the country they live in have high GDP, strong social support, and they could be free to make life choices, also the country is generous and offering positive affect. On the contrary, if a country has a corrupt government, people wouldn’t be happy.

<br>

### Plots of Factors and Happiness score

```{r}
plot_happy = function(A,B,name){
    name = name %>% 
    str_replace_all("_"," ") %>% 
    str_to_title()
    
happy_B = bind_cols(A=A,B=B,country_name = happy_df_fac$country_name,
                    year = happy_df_fac$year) %>% 
  filter(!is.na(A)) %>%
  filter(!is.na(B)) %>%
  arrange(B) %>%
  mutate(
  text_label = 
  str_c("Happiness score: ", A,
        "\n",name,": ", B,
         "\nCountry: ", country_name,
         "\nYear: ", year))

fit_B = lm(A ~ B, data = happy_B)

happy_B%>%
  plot_ly(
    x = ~B,
    y = ~A,
    type = "scatter",
    mode = "markers",
    color = ~year,
    text = ~text_label,
    alpha = 0.5,
    colors = "viridis") %>%
  add_lines(x = ~B, y = predict(fit_B), color = I("red")) %>%
  layout(
    xaxis = list(title = name), yaxis = list(title = "Happiness score")
  )

}
```

<br>

#### The relationship between GDP and Happiness score

```{r}
plot_log = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$log_gdp_per_capita, name = "log_gdp_per_capita")
plot_log
 
log_lm_df = lm(happiness_score ~ log_gdp_per_capita, data = happy_df_fac)

summ(log_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, high GDP has a positive effect on the happiness score. We can see from the table that the p-value equals to 0, which suggests that changes in the GDP are associated with changes in the happiness score. 

<br>

#### The relationship between Social support and Happiness score

```{r}
plot_social = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$social_support, name = "social_support")
plot_social
 
social_lm_df = lm(happiness_score ~ social_support , data = happy_df_fac)

summ(social_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, strong social support has a positive effect on the happiness score. Also, it seems that most countries are offering strong social support since the dots in the plots are concentrated on the right side.

<br>

#### The relationship between Healthy life expectancy at birth and Happiness score

```{r}
plot_healthy = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$healthy_life_expectancy_at_birth, name = "healthy_life_expectancy_at_birth")
plot_healthy
 
healthy_lm_df = lm(happiness_score ~ healthy_life_expectancy_at_birth , data = happy_df_fac)

summ(healthy_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, good healthy life expectancy at birth has a positive effect on the happiness score. We can see from the table that the p-value equals to 0, which suggests that changes in the healthy life expectancy are associated with changes in the happiness score. 

<br>

#### The relationship between Freedom to make life choices and Happiness score

```{r echo = FALSE}
plot_freedom = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$freedom_to_make_life_choices, name = "freedom_to_make_life_choices")
plot_freedom
 
freedom_lm_df = lm(happiness_score ~ freedom_to_make_life_choices , data = happy_df_fac)

summ(freedom_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, if the people are free to make life choices, they are tending to be happy. The p-value equals to 0. We can see in the plot that the dots are widely distributed on both sides around the fitting line, which suggests that the relationship between variables is weak.

<br>

#### The relationship between Generosity and Happiness score

```{r}
plot_generosity = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$generosity, name = "generosity")
plot_generosity
 
generosity_lm_df = lm(happiness_score ~ generosity , data = happy_df_fac)

summ(generosity_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, if the country is generous to the people, the people are tending to be happy. The p-value equals to 0. We can see in the plot that the dots are also widely distributed on both sides around the fitting line, which suggests that the relationship between variables is weak.

<br>

#### The relationship between Corruption and Happiness score

```{r}
plot_perceptions = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$perceptions_of_corruption, name = "perceptions_of_corruption")
plot_perceptions
 
perceptions_lm_df = lm(happiness_score ~ perceptions_of_corruption , data = happy_df_fac)

summ(perceptions_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

Corruption has a clear negative affect on the happiness score. As we can see in the plot, most dots are concentrating on the right side, which means many governments have high scores on corruption.

<br>

#### The relationship between Positive affect and Happiness score

```{r}
plot_positive = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$positive_affect, name = "positive_affect")
plot_positive
 
positive_lm_df = lm(happiness_score ~ positive_affect , data = happy_df_fac)

summ(positive_lm_df)%>%
  tidy() %>%
  knitr::kable()
```

<br>

As shown in the plot, positive affect could make people feel happy. We can see from the table that the p-value equals to 0, which suggests that changes in the positive affect are associated with changes in the happiness score. 

<br>

#### The relationship between Negative affect and Happiness score

```{r}
plot_negative = plot_happy(A=happy_df_fac$happiness_score,B =happy_df_fac$negative_affect, name = "negative_affect")
plot_negative
 
negative_lm_df = lm(happiness_score ~ negative_affect , data = happy_df_fac)

summ(negative_lm_df)%>%
  tidy() %>%
  knitr::kable()

```

<br>

On the contrary, negative affect is leading to low happiness score. We can see from the table that the p-value equals to 0, which suggests that changes in the negative affect are associated with changes in the happiness score. Also, most dots are concentrating on the left, which means few countries are offering negative score. 
